---
title: "191011_OP56_ELT2_vs_L4440_RNAi_Analysis"
author: "RTPW"
date: "10/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries

```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(ggpubr)
```


Load in data

```{r}
image_df_1 <- read_csv("190626_ELT-2_ImageJ_Analysis.csv", col_names = TRUE)
image_df_1$RNAi_Treatment <- factor(image_df$RNAi_Treatment, levels = c("L4440", "ELT-2"))
image_df_1$Rep <- 1
```


```{r}
image_df_2 <- read_csv("201130_OP56_ELT-2_RNAi_Analysis.csv", col_names = TRUE)
head(image_df_2)
image_df <- image_df_1 %>% bind_rows(image_df_2)
```

Calculate the corrected total cell fluorescence (CTCF)

CTCF = Integrated Density – (Area of selected cell X Mean fluorescence of background readings)

Calculate the mean background value for each image
```{r}
background_df<- image_df %>%
  group_by(RNAi_Treatment, Worm_Num, Measurement_Type) %>%
  summarize(Mean_Background = mean(Mean_Grey_Value)) %>% 
  filter(Measurement_Type == "Background")
background_df
```

Combine the mean of mean background values with the integrated density and intestine area. Calculate the corrected total cell flourescence.

```{r}
image_ctcf <- image_df %>%
  filter(Measurement_Type == "Worm") %>%
  select(RNAi_Treatment, Worm_Num, Measurement_Area, Integrated_Density) %>%
  mutate(Mean_Background = background_df$Mean_Background) %>%
  mutate(CTCF = Integrated_Density - (Measurement_Area*Mean_Background) )
image_ctcf  
```

Plot the results

```{r}
ggplot(image_ctcf, aes(RNAi_Treatment, CTCF)) + 
  geom_boxplot() +
  geom_jitter() +
  labs(
    title = "ELT-2 RNAi Knockdown \nDecreases ELT-2::GFP Abundance",
    x = "RNAi Treatment",
    y = "Corrected Total Cell \nFluorescence (A.U.)"
  ) +
  stat_compare_means(comparisons = list(c("L4440", "ELT-2")), method = "t.test") +
  scale_x_discrete(labels = c("Empty Vector \n(L4440)", "ELT-2\n(pRW17)"))+
  theme_classic() +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1),
    axis.text.x = element_text(size = 18),
    axis.text.y = element_text(size = 16),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size=18),
    plot.title = element_text(hjust = 0.5, size = 20)
    )
  
```

Save the plot

```{r}
ggsave("191011_OP56_ELT2_vs_L4440_RNAi_GFP_Quant_Plot.png", width = 5, height = 5.15)
```



