---
title: "RWP32_ELT2gfp_RNAi_Analysis"
author: "RTPW"
date: "8/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries

```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(tidyverse)
```


Load in data

```{r}
# for rep 1, CH1 = GFP and CH2 = DAPI (gut granules)
image_df_rep1 <- read_csv("./01_input/ELT2gfp_REP1_quantification.csv", col_names = TRUE)%>% separate(col = Label, sep = ":", into = c("treatment", "measurement_type", "image_name")) %>% separate(col = image_name, sep = "_", into = c("max", "strain", "Treatment", "worm", "wormNum", "channel"), remove = FALSE) %>% dplyr::select(image_name, treatment, wormNum, measurement_type, channel, Area:RawIntDen)
image_df_rep1$Rep <- 1
# for rep 2 and rep 3, CH2 = GFP and CH3 = DAPI (gut granules)
# use mutate to change CH2 to CH1 and CH3 to CH2
image_df_rep2 <- read_csv("./01_input/ELT2gfp_REP2_quantification.csv", col_names = TRUE) %>% separate(col = Label, sep = ":", into = c("treatment", "measurement_type", "image_name")) %>% separate(col = image_name, sep = "_", into = c("max", "strain", "Treatment", "worm", "wormNum", "channel"), remove = FALSE) %>% dplyr::select(image_name, treatment, wormNum, measurement_type, channel, Area:RawIntDen) %>% mutate(case_when(channel == "C2" ~ "C1", channel == "C3" ~ "C2"))
image_df_rep2$Rep <- 2

image_df_rep3 <- read_csv("./01_input/ELT2gfp_REP3_quantification.csv", col_names = TRUE) %>% separate(col = Label, sep = ":", into = c("treatment", "measurement_type", "image_name")) %>% separate(col = image_name, sep = "_", into = c("max", "strain", "Treatment", "worm", "wormNum", "channel"), remove = FALSE) %>% dplyr::select(image_name, treatment, wormNum, measurement_type, channel, Area:RawIntDen) %>% mutate(channel = case_when(channel == "C2" ~ "C1", channel == "C3" ~ "C2"))
image_df_rep3$Rep <- 3

image_df <- bind_rows(image_df_rep1, image_df_rep2,image_df_rep3)
image_df$treatment <- factor(image_df$treatment, levels = c("L4440rnai", "ELT2rnai"))
head(image_df)

```

Calculate the corrected total cell fluorescence (CTCF)

CTCF = Integrated Density – (Area of selected cell X Mean fluorescence of background readings)

Calculate the mean background value for each image
```{r}
background_df<- image_df %>%
  group_by(treatment, wormNum, Rep, measurement_type, channel) %>%
  summarize(Mean_Background = mean(Mean)) %>% 
  filter(measurement_type == "background")
background_df
```

Combine the mean of mean background values with the integrated density and intestine area. Calculate the corrected total cell flourescence.

```{r}
image_ctcf <- image_df %>%
  filter(measurement_type == "intestine") %>%
  select(treatment, wormNum, Rep, channel, Area, IntDen) %>%
  mutate(Mean_Background = background_df$Mean_Background) %>%
  mutate(CTCF = IntDen - (Area*Mean_Background) )
image_ctcf  
```

Plot the results

```{r fig.height=4, fig.width=4}
image_ctcf %>% filter(channel == "C1") %>%
ggplot(aes(treatment, CTCF/Area)) + 
  geom_boxplot(width = 0.5) +
  geom_jitter(width = 0.25, alpha = 0.5, shape = 16) +
  labs(y = "intestine fluorescence (A.U.)", x = NULL) +
  scale_x_discrete(labels = c("control RNAi", "elt-2 RNAi"))+
  theme_classic() +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size=18),
    plot.title = element_text(hjust = 0.5, size = 18)
    )
# +
#   stat_compare_means(comparisons = list(c("L4440", "ELT-2")), method = "t.test") +
#   scale_x_discrete(labels = c("Empty Vector \n(L4440)", "ELT-2\n(pRW17)"))+
#   theme_classic() +
#   theme(
#     panel.border = element_rect(colour = "black", fill = NA, size = 1),
#     axis.text.x = element_text(size = 18),
#     axis.text.y = element_text(size = 16),
#     axis.title.x = element_text(size = 20),
#     axis.title.y = element_text(size=18),
#     plot.title = element_text(hjust = 0.5, size = 20)
#     )

ggsave("./03_plots/RWP32_ELT2gfp_Quantification_Plot.pdf", width = 4, height = 4, useDingbats=FALSE)
```
```{r}
gfp_elt2 <- (image_ctcf %>% filter(treatment == "ELT2rnai") %>% rowwise() %>% mutate(normGFP = CTCF/Area))$normGFP
gfp_L4440 <- (image_ctcf %>% filter(treatment == "L4440rnai") %>% rowwise() %>% mutate(normGFP = CTCF/Area))$normGFP

t.test(gfp_elt2, gfp_L4440)$p.value

image_ctcf %>% group_by(treatment) %>% summarise(n())
```

```{r}
sessionInfo()
```

