---
title: "CEBP1gfp_Image_Analysis"
author: "Robert Williams"
date: "6/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install Packages
Install the necessary packages. Do this once the first time analysis is performed.   Uncomment and execute the following code chunk.

```{r}
# install.packages("tidyverse")
# install.packages("readxl")
# install.packages("ggpubr")
# install.packages("mratios")
```


## Load Package Libraries

```{r echo=FALSE}
library(tidyverse)
library(readxl)
library(ggpubr)
library(mratios)
```

# Load in quantification data

Each worm should have the following measurements:

- One intestine fluorescence measurement
- Four background measurements

The input data should have the following columns (without quotes):

 - "treatment": The type of RNAi treatment
 - "life_stage": The worm developmental stage
 - "worm": the worm number imaged
 - "measurement_type": `background` or `GFP`
 - "measurement_num": the measurement number from ImageJ
 - "area": region of interest area
 - "mean": mean gray pixel intensity
 - "min": minimum pixel intensity
 - "max": maximum pixel intensity
 - "intDen": product of area and sum of the values of the pixels in the selection
 - "rawIntDen": sum of the values of the pixels in the selection
 - "experiment": experiment name including rep number
 - "strain": strain or reporter gene name
 
More info here: https://imagej.nih.gov/ij/docs/guide/146-30.html

Below is code that will work for the data in the `example_data` directory. This section may need to be changed depending on how the input data is structured.

```{r}
image_df <- read_csv(file = "./01_input/CEBP1gfp_L4440rnai_Rep3_Quant.csv") %>% 
  bind_rows(read_csv(file = "./01_input/CEBP1gfp_ELT2rnai_Rep3_Quant.csv"))

image_df_rep3 <- image_df %>% separate(col = Label, sep = ":", into = c("image_name", "measurement_type", "channel")) %>% separate(col = image_name, sep = "_", into = c("max", "strain", "treatment", "Rep", "Obj", "worm", "Extra"), remove = FALSE) %>% separate(measurement_type, sep = "\\d+", into = "measurement_type") %>% 
  dplyr::select(image_name, treatment, measurement_type, channel, Area:RawIntDen)
image_df_rep3$channel <- factor(image_df_rep3$channel, levels = c(2, 1), labels = c("C1", "C2"))
image_df_rep3$treatment <- factor(image_df_rep3$treatment, levels = c("L4440rnai", "ELT2rnai"), labels = c("01_L4440rnai", "03_ELT2rnai"))
image_df_rep3$rep <- "Rep3"

image_df_rep4 <- read_csv(file = "./01_input/210428_CEBP1_Rep4_Quantification.csv") %>% separate(Label, c("treatment", "measurement_type", "image_name"), ":") %>% separate(image_name, into = c("image_name_2","channel"), -2, remove = FALSE) %>%
  dplyr::select(image_name, treatment, measurement_type, channel, Area:RawIntDen)
image_df_rep4$rep <- "Rep4"

image_df_rep5 <- read_csv(file = "./01_input/210501_CEBP1gfp_Rep5_Quantification.csv") %>% separate(Label, c("treatment", "measurement_type", "image_name"), ":") %>% separate(image_name, into = c("image_name_2","channel"), -2, remove = FALSE) %>%
  dplyr::select(image_name, treatment, measurement_type, channel, Area:RawIntDen)
image_df_rep5$rep <- "Rep5"



image_df <- bind_rows(image_df_rep3, image_df_rep4, image_df_rep5)

# Set the factor levels for ordering in downstream plotting
image_df$treatment <- factor(image_df$treatment, levels = c("01_L4440rnai", "02_ELT7rnai", "03_ELT2rnai", "04_ELT2ELT7rnai"))
image_df <- image_df %>% mutate(image_name = paste(image_name, channel, rep, sep = "_"))
nrow(image_df)
```


# Calculate corrected total cell fluorescence (CTCF)

CTCF = Integrated Density – (Area of selected cell X Mean fluorescence of background readings)  
More information here: https://theolb.readthedocs.io/en/latest/imaging/measuring-cell-fluorescence-using-imagej.html

```{r}
# Take the mean of the background measurements
background_df<- image_df %>%
  group_by(treatment, image_name, measurement_type, channel, rep) %>%
  filter(measurement_type == "background", channel == "C2") %>%
  summarize(Mean_Background = mean(Mean)) %>% 
  ungroup() %>%
  dplyr::select(image_name, Mean_Background, channel, rep)
background_df
```
```{r}
# calculate CTCF for each worm
image_ctcf <- image_df %>%
  filter(measurement_type %in% c("intestine","worm"), channel == "C2") %>%
  dplyr::select(treatment, image_name, Area, IntDen) %>%
  left_join(background_df, by = "image_name") %>%
  rowwise() %>%
  mutate(CTCF = IntDen - (Area*Mean_Background) )
image_ctcf
```

# Plot the quantification results

```{r fig.height=4, fig.width=4}

image_df %>% filter(channel == "C2", measurement_type == "background") %>% group_by(image_name, measurement_type) %>% summarise(Mean_Background = mean(Mean)) %>% dplyr::select(-measurement_type) %>% right_join(image_df, by = "image_name") %>% filter(measurement_type %in% c("worm", "intestine"), treatment %in% c("01_L4440rnai", "03_ELT2rnai")) %>% mutate(CTCF = IntDen - (Area*Mean_Background)) %>% filter(CTCF > 1, Area > 300) %>%
  ggplot(aes(treatment, CTCF/Area)) + 
  geom_boxplot(width = 0.5) +
  geom_jitter(width = 0.25, alpha = 0.5, shape = 16, aes(color = rep)) +
  # facet_grid(.~channel) +
  labs(
    title = "CEBP-1::GFP Reporter",
    x = "RNAi Treatment",
    y = "Intestine Fluorescence (A.U.)"
  ) +
  # stat_compare_means(comparisons = list(c("L4440rnai", "ELT7rnai")), method = "t.test") +
  # stat_compare_means(comparisons = list(c("L4440rnai", "ELT2rnai")), method = "t.test") +
  # stat_compare_means(comparisons = list(c("L4440rnai", "ELT2ELT7rnai")), method = "t.test") +
  scale_x_discrete(labels = c("L4440", "ELT-2"))+
  theme_classic() +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size=18),
    plot.title = element_text(hjust = 0.5, size = 18)
    ) 

# ggsave("./RWP32_CEBP1gfp_Quantification_Plot.pdf", width = 5, height = 4, useDingbats=FALSE)
```

```{r}
gfp_elt2 <- (image_ctcf %>% filter(treatment == "03_ELT2rnai") %>% rowwise() %>% mutate(normGFP = CTCF/Area))$normGFP
gfp_L4440 <- (image_ctcf %>% filter(treatment == "01_L4440rnai") %>% rowwise() %>% mutate(normGFP = CTCF/Area))$normGFP

t.test(gfp_elt2, gfp_L4440)$p.value

image_ctcf %>% group_by(treatment) %>% summarise(n())
```


# TO DO
# Measure fold change 
To see how much brighter ELT-2 is compared to L4440

```{r}
mean_ctcf <-
  image_ctcf %>%
  group_by(strain, treatment) %>%
  summarise(avgCTCF = mean(CTCF), sdCTCF = sd(CTCF))

mean_ctcf <-
  mean_ctcf %>% pivot_wider(names_from = treatment,
                            values_from = c(avgCTCF, sdCTCF))
mean_ctcf
```
# Calculate fold change and add confidence intervals

```{r}

image_ctcf %>%filter(life_stage=='L1' & treatment == 'ELT-2', strain == "JM149") %>% dplyr::select(CTCF) -> JM149_numerator
image_ctcf %>% filter(life_stage=='L1' & treatment == 'L4440', strain == "JM149") %>% dplyr::select(CTCF) -> JM149_denominator
ttestratio(JM149_numerator[[1]], JM149_denominator[[1]]) -> L1_ttest_ratio_JM149

image_ctcf %>%filter(life_stage=='L1' & treatment == 'ELT-2', strain == "JM259") %>% dplyr::select(CTCF) -> JM259_numerator
image_ctcf %>% filter(life_stage=='L1' & treatment == 'L4440', strain == "JM259") %>% dplyr::select(CTCF) -> JM259_denominator
ttestratio(JM259_numerator[[1]], JM259_denominator[[1]]) -> L1_ttest_ratio_JM259
    
mean_ctcf$lower = c(L1_ttest_ratio_JM149$conf.int[1], L1_ttest_ratio_JM259$conf.int[1])
mean_ctcf$upper = c(L1_ttest_ratio_JM149$conf.int[2], L1_ttest_ratio_JM259$conf.int[2])
```

# Plot fold change results
```{r fig.width=3, fig.height=3}
ggplot(mean_ctcf, aes(x = strain, y = `avgCTCF_ELT-2`/avgCTCF_L4440)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  labs(title = "elt-2 Promoter\nRelative Fluorescence") +
  theme_classic()+
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 0.5),
    axis.text.x = element_text(size = 16),
    axis.text.y = element_text(size = 16),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size=20),
    plot.title = element_text(hjust = 0.5, size = 20)
    ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  ylab("Relative Fluorescence") + coord_flip()

# ggsave("./ELT-2_promoter_analysis_fold_change_210117.pdf", width = 3, height = 3, useDingbats=FALSE)  
```
