---
title: "RWP32_ELT2gfp_RNAi_Analysis"
author: "RTPW"
date: "8/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries

```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(ggpubr)
```


Load in data

```{r}
image_df_1 <- read_csv("./01_input/ELT2gfp_REP1_quantification.csv", col_names = TRUE)

image_df <- image_df_1 %>% separate(col = Label, sep = ":", into = c("treatment", "measurement_type", "image_name")) %>% separate(col = image_name, sep = "_", into = c("max", "strain", "Treatment", "worm", "wormNum", "channel"), remove = FALSE) %>% 
  dplyr::select(image_name, treatment, wormNum, measurement_type, channel, Area:RawIntDen)
image_df$treatment <- factor(image_df$treatment, levels = c("L4440rnai", "ELT2rnai"))
image_df$Rep <- 1
head(image_df)
```

Calculate the corrected total cell fluorescence (CTCF)

CTCF = Integrated Density – (Area of selected cell X Mean fluorescence of background readings)

Calculate the mean background value for each image
```{r}
background_df<- image_df %>%
  group_by(treatment, wormNum, measurement_type, channel) %>%
  summarize(Mean_Background = mean(Mean)) %>% 
  filter(measurement_type == "background")
background_df
```

Combine the mean of mean background values with the integrated density and intestine area. Calculate the corrected total cell flourescence.

```{r}
image_ctcf <- image_df %>%
  filter(measurement_type == "intestine") %>%
  select(treatment, wormNum, channel, Area, IntDen) %>%
  mutate(Mean_Background = background_df$Mean_Background) %>%
  mutate(CTCF = IntDen - (Area*Mean_Background) )
image_ctcf  
```

Plot the results

```{r fig.width = 4, fig.height= 4}
image_ctcf %>% filter(channel == "C1") %>%
ggplot(aes(treatment, CTCF/Area)) + 
  geom_boxplot(width = 0.5) +
  geom_jitter(width = 0.25, alpha = 0.5, shape = 16) +
  labs(y = "intestine fluorescence (A.U.)", x = NULL) +
  scale_x_discrete(labels = c("control RNAi", "elt-2 RNAi"))+
  theme_classic() +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size=18),
    plot.title = element_text(hjust = 0.5, size = 18)
    )
# +
#   stat_compare_means(comparisons = list(c("L4440", "ELT-2")), method = "t.test") +
#   scale_x_discrete(labels = c("Empty Vector \n(L4440)", "ELT-2\n(pRW17)"))+
#   theme_classic() +
#   theme(
#     panel.border = element_rect(colour = "black", fill = NA, size = 1),
#     axis.text.x = element_text(size = 18),
#     axis.text.y = element_text(size = 16),
#     axis.title.x = element_text(size = 20),
#     axis.title.y = element_text(size=18),
#     plot.title = element_text(hjust = 0.5, size = 20)
#     )

# ggsave("./03_plots/RWP32_ELT2gfp_Quantification_Plot.pdf", width = 4, height = 4, useDingbats=FALSE)
```

