---
title: "KW_ELT2_Figure3"
author: "Robert Williams"
date: "3/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(biomaRt)
library(tidyverse)
library(readxl)
library(openxlsx)
library(ComplexHeatmap)
library(lubridate)
```

Load custom functions

```{r}
source("../RWC23_Functions.R")
```

Set plot variable

```{r}
plot <- FALSE
plotdir <- "./03_plots/"
```


# Import Time-resolved RNA

```{r}
time_resolved_rna <- read.delim("./01_input/9_Boeck_et_al_2016_time-resolved_transcriptome/Unified_dcpm_per_wormbase_gene.txt", quote="", stringsAsFactors=FALSE)

paramart <- useMart("parasite_mart", dataset = "wbps_gene", host = "https://parasite.wormbase.org", port = 443)

time_resolved_rna <- getBM(
  mart = paramart,
  filter = c("wormbase_gseqname"),
  value = time_resolved_rna$WormbaseName,
  attributes = c("wormbase_gseq","wbps_gene_id", "wikigene_name")
) %>% right_join(time_resolved_rna, by = c("wormbase_gseq" = "WormbaseName"))


time_resolved_rna <- time_resolved_rna %>% drop_na(wbps_gene_id)
```


# Load in public intestine expression datasets

Initalize an empty dataframe.

```{r}
intestine_rna <- data.frame(WBGeneID = time_resolved_rna$wbps_gene_id, Sequence.name = time_resolved_rna$wormbase_gseq, Public_name = time_resolved_rna$wikigene_name)
```


# A Transcriptional Lineage of the Early C . elegans Embryo.
## Tintori SC, Nishimura EO, Golden P, Lieb JD, Goldstein B. Dev Cell. 2016;38(4):430-444. doi:10.1101/047746

```{r}
tintori_2Ep <- read.csv("./01_input/7_Tintori_et_al_2016_Embryo_scRNA-seq/16cell_vs_Ep_p1_00.csv", header = FALSE, stringsAsFactors = FALSE)
tintori_columns <- c("gene", "p.value", "log.CPM", "log.FC")
colnames(tintori_2Ep) <- str_c("tintori_2Ep_", tintori_columns)
tintori_2Ep <- tintori_2Ep %>% mutate(tintori_2Ep_log.FC = tintori_2Ep_log.FC*-1)
intestine_rna <- intestine_rna %>% left_join(tintori_2Ep, by = c("Sequence.name" = "tintori_2Ep_gene"))

```

# Spatiotemporal transcriptomics reveals the evolutionary history of the endoderm germ layer.
## Hashimshony T, Feder M, Levin M, Hall BK, Yanai I.  Nature. 2015;519(7542):219-222. doi:10.1038/nature13996

```{r}
hashimshony_embryo <- read.csv("./01_input/1_Hashimshony_et_al_2015_CEL-seq/Hashimshony_Endoderm_Genes.csv", header = TRUE)
colnames(hashimshony_embryo) <- c("hashimshony_embryo_genes")

```

# Chromosomal clustering and GATA transcriptional regulation of intestine-expressed genes in C. elegans.
## Pauli F, Liu Y, Kim Y a, Chen P-J, Kim SK. Development. 2006;133(2):287-295. doi:10.1242/dev.02185

```{r}
# TFs from Pauli et al 
pauliGenes <- read_excel("./01_input/4_Pauli_et_al_2006_PAB-1_Pulldown_microarray/SupTable1_intestine_enriched_genes.xlsx")
colnames(pauliGenes) <- str_c("pauli_L4_", c("gene", "percentile", "pvalue", "description"))
# add to master datatable
intestine_rna <- intestine_rna %>% left_join(pauliGenes, by = c("Sequence.name" = "pauli_L4_gene"))
```

# A spatial and temporal map of C. elegans gene expression. Genome Research, 21(2), 325–41. https://doi.org/10.1101/gr.114595.110 
## Spencer, W. C., Zeller, G., Watson, J. D., Henz, S. R., Watkins, K. L., McWhirter, R. D., … Miller, D. M. (2011).

```{r}

# Late embryo intestine from Spencer et al
spencerLEgenes <- read.table("./01_input/6_Spencer_et_al_2010_FACS_and_pulldown_tilling_array/LE-intestine_enr_vs_ref.WS200.txt", quote="\"", comment.char="", header = TRUE)
colnames(spencerLEgenes) <- str_c("spencer_LE_", colnames(spencerLEgenes))
spencer_LE_subset <- spencerLEgenes %>% dplyr::select(spencer_LE_ID, spencer_LE_AveExpr, spencer_LE_adj_P_Val, spencer_LE_FC)

# add to dataframe
intestine_rna <- merge(x = intestine_rna, y = spencer_LE_subset, by.x = 'WBGeneID', by.y = 'spencer_LE_ID', all.x = TRUE)



# L2 intestine genes from Spencer et al
spencerL2genes <- read.table("./01_input/6_Spencer_et_al_2010_FACS_and_pulldown_tilling_array/L2-intestine_enr_vs_ref.WS200.txt", quote="\"", comment.char="", header = TRUE)
colnames(spencerL2genes) <- str_c("spencer_L2_", colnames(spencerL2genes))
spencer_L2_subset <- spencerL2genes %>% dplyr::select(spencer_L2_ID, spencer_L2_AveExpr, spencer_L2_adj_P_Val, spencer_L2_FC)

intestine_rna <- merge(x = intestine_rna, y = spencer_L2_subset, by.x = 'WBGeneID', by.y = 'spencer_L2_ID', all.x = TRUE)

```

# Final Filtering

```{r}
# Tintori
intestine_rna <-
  intestine_rna %>% mutate(
    tintori_2Ep_bool = if_else(
      intestine_rna$Sequence.name %in%
        (tintori_2Ep %>% filter(tintori_2Ep_log.FC > 1.75 & tintori_2Ep_p.value <= 0.05))$tintori_2Ep_gene,
      TRUE,
      FALSE
    )
  )
nrow(intestine_rna %>% filter(tintori_2Ep_bool == TRUE))
# Hashimshony
intestine_rna <- intestine_rna %>% mutate(hashimshony_embryo_bool = if_else(intestine_rna$Public_name %in% hashimshony_embryo$hashimshony_embryo_genes, TRUE, FALSE))
nrow(intestine_rna %>% filter(hashimshony_embryo_bool == TRUE))

# Pauli
# intestine_rna <- intestine_rna %>% mutate(pauli_L4_bool = if_else(intestine_rna$Sequence.name %in% (pauliGenes %>% filter(pauli_L4_percentile > 0.8))$pauli_L4_gene, TRUE, FALSE))
intestine_rna <-
  intestine_rna %>% mutate(pauli_L4_bool = if_else(
    intestine_rna$Sequence.name %in% pauliGenes$pauli_L4_gene,
    TRUE,
    FALSE
  ))
nrow(intestine_rna %>% filter(pauli_L4_bool == TRUE))

# Spencer Late Embryo
intestine_rna <- intestine_rna %>% mutate(spencer_LE_bool = if_else(intestine_rna$WBGeneID %in% spencer_LE_subset$spencer_LE_ID, TRUE, FALSE))
nrow(intestine_rna %>% filter(spencer_LE_bool == TRUE))

# Spencer L2
intestine_rna <- intestine_rna %>% mutate(spencer_L2_bool = if_else(intestine_rna$WBGeneID %in% spencer_L2_subset$spencer_L2_ID, TRUE, FALSE))
nrow(intestine_rna %>% filter(spencer_L2_bool == TRUE))
```
```{r}
intestine_rna %>% filter(Public_name %in% c("M88.4", "bcl-11", "sre-23", "T25D10.4", "aat-6", "pgp-1", "ftn-1", "elt-2", "elt-7")) %>% dplyr::select(Public_name, contains("bool"))
```


Make a vector that contains intestine expressed genes from any stage

```{r}
intestine_gene_list <- intestine_rna %>% filter(
  tintori_2Ep_bool == TRUE |
    hashimshony_embryo_bool == TRUE |
    pauli_L4_bool == TRUE |
    spencer_LE_bool == TRUE |
    spencer_L2_bool == TRUE
) %>% dplyr::select(WBGeneID, Public_name, contains("bool")) %>% dplyr::rename(
  tintori_2Ep = "tintori_2Ep_bool",
  hashimshony_emb = "hashimshony_embryo_bool",
  pauli_L4 = "pauli_L4_bool",
  spencer_LE = "spencer_LE_bool",
  spencer_L2 = "spencer_L2_bool"
)

head(intestine_gene_list)
dim(intestine_gene_list)
```

Write output files

```{r}
# write_csv(intestine_gene_list, "./02_output/Fig3_Public_Intestine_RNA_Gene_List.csv")
# write_csv(intestine_rna, "./02_output/Fig3_Public_Intestine_RNA_Full_Data.csv")
```

Read in files for downstream work

```{r}
# intestine_rna <- read.csv("./02_output/Fig3_Public_Intestine_RNA_Full_Data.csv", stringsAsFactors = FALSE)
# intestine_gene_list <- read.csv("./02_output/Fig3_Public_Intestine_RNA_Gene_List.csv", stringsAsFactors = FALSE)
```


# UpSet plot of overlaps

```{r}
# there are 5 sets of interest
# set up the combination matrix
intestine_upset <-
  intestine_gene_list %>% left_join(intestine_rna, by = "WBGeneID") %>% dplyr::select(WBGeneID, contains("bool")) %>% column_to_rownames(var = "WBGeneID") %>% dplyr::rename(
    tint_2Ep = "tintori_2Ep_bool",
    hash_emb = "hashimshony_embryo_bool",
    pauli_L4 = "pauli_L4_bool",
    spen_LE = "spencer_LE_bool",
    spen_L2 = "spencer_L2_bool"
  )
head(intestine_upset)
nrow(intestine_upset)
# replace TRUE/FALSE with 1/0
intestine_upset[, sapply(intestine_upset, is.logical)] <-
  lapply(intestine_upset[, sapply(intestine_upset, is.logical)], as.numeric)
# convert to matrix
intestine_upset_matrix <- as.matrix.data.frame(intestine_upset)
# peek
head(intestine_upset_matrix)
# number of rows
nrow(intestine_upset_matrix)
```

```{r}
m1 <- make_comb_mat(intestine_upset_matrix)
m1
```

```{r fig.width=5, fig.height=3}
intestine_upset_plot <- UpSet(m1, 
      set_order = c("tint_2Ep", "hash_emb", "spen_LE", "spen_L2", "pauli_L4"),
      comb_order = order(comb_size(m1)))
intestine_upset_plot

if(plot == TRUE){
  myPDFplot(intestine_upset_plot, name = "Intestine_Gene_Upset_Plot", height = 3, width = 5,plotdir)
}
```


# Metadata

Initalize a seprate empty "metadata" dataframe to store descriptive information

```{r}
RWC23_Metadata <- data.frame(column_ID = "", column_detail = "", source_DOI = "", source_detail = "", seq_method = "")

# Tintori metadata

RWC23_Metadata <- RWC23_Metadata %>% add_row(
  column_ID = colnames(tintori_2Ep[1]),
  column_detail = "Gene sequence name",
  source_DOI = "10.1016/j.devcel.2016.07.025",
  source_detail = "Single cell RNA-seq of hand dissected embryo. Accessed from tintori.bio.unc.edu.",
  seq_method = "Illumina"
  ) %>% add_row(
  column_ID = colnames(tintori_2Ep[2]),
  column_detail = "Gene p-value",
  source_DOI = "10.1016/j.devcel.2016.07.025",
  source_detail = "Single cell RNA-seq of hand dissected embryo. Accessed from tintori.bio.unc.edu.",
  seq_method = "Illumina"
  ) %>% add_row(
  column_ID = colnames(tintori_2Ep[3]),
  column_detail = "Counts per transcript per million mapped reads.",
  source_DOI = "10.1016/j.devcel.2016.07.025",
  source_detail = "Single cell RNA-seq of hand dissected embryo. Accessed from tintori.bio.unc.edu.",
  seq_method = "Illumina"
  ) %>% add_row(
  column_ID = colnames(tintori_2Ep[4]),
  column_detail = "Log2(Fold Change) of gene in the 2Ep cell when compared to the whole embryo. Less than zero is enriched in the 1E cell. Greater than zero is enriched in the whole 16 cell embryo",
  source_DOI = "10.1016/j.devcel.2016.07.025",
  source_detail = "Single cell RNA-seq of hand dissected embryo. Accessed from tintori.bio.unc.edu.",
  seq_method = "Illumina"
  )


# Hashimshony metadata

RWC23_Metadata <- RWC23_Metadata %>% add_row(
  column_ID = colnames(hashimshony_embryo[1]),
  column_detail = "Gene public name",
  source_DOI = "10.1038/nature13996",
  source_detail = "Supplementary Table 4. Endoderm gene set from Single cell RNA-seq of cultured dissociated embryos.",
  seq_method = "Illumina"
  )

# update metadata
RWC23_Metadata <- RWC23_Metadata  %>% add_row(
  column_ID = colnames(pauliGenes[1]),
  column_detail = "Sequence name",
  source_DOI = "10.1242/dev.02185",
  source_detail = "Supplementary Table 1. Microarray of ges-1P::PAB-1 pulldown in L4 worms"
  ) %>% add_row(
  column_ID = colnames(pauliGenes[2]),
  column_detail = "Percentile ranking of transcript abundance",
  source_DOI = "10.1242/dev.02185",
  source_detail = "Supplementary Table 1. Microarray of ges-1P::PAB-1 pulldown in L4 worms"
  ) %>% add_row(
  column_ID = colnames(pauliGenes[3]),
  column_detail = "Student's T-test to calculate significant enrichment of intestine transcripts",
  source_DOI = "10.1242/dev.02185",
  source_detail = "Supplementary Table 1. Microarray of ges-1P::PAB-1 pulldown in L4 worms"
  ) %>% add_row(
  column_ID = colnames(pauliGenes[4]),
  column_detail = "Short descriptions generated by WormBase",
  source_DOI = "10.1242/dev.02185",
  source_detail = "Supplementary Table 1. Microarray of ges-1P::PAB-1 pulldown in L4 worms"
  ) %>% add_row(
  column_ID = "pauli_L4_bool",
  column_detail = "True if included on the list, false if absent on the list",
  source_DOI = "10.1242/dev.02185",
  source_detail = "Supplementary Table 1. Microarray of ges-1P::PAB-1 pulldown in L4 worms"
  ) 

# add sequencing method column

RWC23_Metadata <- mutate(RWC23_Metadata, seq_method = ifelse(RWC23_Metadata$column_ID %in% c(colnames(pauliGenes), "pauli_L4_bool"), "Microarray",RWC23_Metadata$seq_method))


# Add Spencer metadata

RWC23_Metadata <- RWC23_Metadata  %>% add_row(
  column_ID = colnames(spencer_LE_subset[1]),
  column_detail = "Wormbase Gene ID",
  source_DOI = "10.1101/gr.114595.110",
  source_detail = "FACS Sorted ELT-2::GFP worms. https://www.vanderbilt.edu/wormdoc/wormmap"
  ) %>% add_row(
  column_ID = colnames(spencer_LE_subset[2]),
  column_detail = "??",
  source_DOI = "10.1101/gr.114595.110",
  source_detail = "FACS Sorted ELT-2::GFP worms. https://www.vanderbilt.edu/wormdoc/wormmap"
  ) %>% add_row(
  column_ID = colnames(spencer_LE_subset[3]),
  column_detail = "??",
  source_DOI = "10.1101/gr.114595.110",
  source_detail = "FACS Sorted ELT-2::GFP worms. https://www.vanderbilt.edu/wormdoc/wormmap"
  ) %>% add_row(
  column_ID = colnames(spencer_LE_subset[4]),
  column_detail = "??",
  source_DOI = "10.1101/gr.114595.110",
  source_detail = "FACS Sorted ELT-2::GFP worms. https://www.vanderbilt.edu/wormdoc/wormmap"
  ) %>% add_row(
  column_ID = "spencer_LE_bool",
  column_detail = "True if present in this list",
  source_DOI = "10.1101/gr.114595.110",
  source_detail = "FACS Sorted ELT-2::GFP worms. https://www.vanderbilt.edu/wormdoc/wormmap"
  )

RWC23_Metadata <- mutate(RWC23_Metadata, seq_method = ifelse(RWC23_Metadata$column_ID %in% c(colnames(spencer_LE_subset), "spencer_LE_bool"), "Microarray",RWC23_Metadata$seq_method))

RWC23_Metadata <- RWC23_Metadata  %>% add_row(
  column_ID = colnames(spencer_L2_subset[1]),
  column_detail = "Wormbase Gene ID",
  source_DOI = "10.1101/gr.114595.110",
  source_detail = "FACS Sorted ELT-2::GFP worms. https://www.vanderbilt.edu/wormdoc/wormmap"
  ) %>% add_row(
  column_ID = colnames(spencer_L2_subset[2]),
  column_detail = "??",
  source_DOI = "10.1101/gr.114595.110",
  source_detail = "FACS Sorted ELT-2::GFP worms. https://www.vanderbilt.edu/wormdoc/wormmap"
  ) %>% add_row(
  column_ID = colnames(spencer_L2_subset[3]),
  column_detail = "??",
  source_DOI = "10.1101/gr.114595.110",
  source_detail = "FACS Sorted ELT-2::GFP worms. https://www.vanderbilt.edu/wormdoc/wormmap"
  ) %>% add_row(
  column_ID = colnames(spencer_L2_subset[4]),
  column_detail = "??",
  source_DOI = "10.1101/gr.114595.110",
  source_detail = "FACS Sorted ELT-2::GFP worms. https://www.vanderbilt.edu/wormdoc/wormmap"
  ) %>% add_row(
  column_ID = "spencer_L2_bool",
  column_detail = "True if present in this list",
  source_DOI = "10.1101/gr.114595.110",
  source_detail = "FACS Sorted ELT-2::GFP worms. https://www.vanderbilt.edu/wormdoc/wormmap"
  )

RWC23_Metadata <- mutate(RWC23_Metadata, seq_method = ifelse(RWC23_Metadata$column_ID %in% c(colnames(spencer_L2_subset), "spencer_L2_bool"), "Microarray",RWC23_Metadata$seq_method))
```

## write output metadata file

```{r}
# write.xlsx(RWC23_Metadata, "./02_output/Fig3_Public_Intestine_RNA_Metadata.xlsx")
```





# Import ELT-2 ChIP-seq binding data

Make sure annotatedPeaks file is linked to `01_input` directory.  
The path is: `../../../ELT2_binding/annotatedPeaks.rds`

```{r}
elt2_GRange <- readRDS("01_input/annotatedPeaks.rds")
head(elt2_GRange)
elt2_peaks <- as.data.frame(mcols(elt2_GRange))
elt2_peaks <- elt2_peaks %>% dplyr::rename(cluster.description = k4labels, WBGeneID = feature)
elt2_cluster_names <- c("Embryo_Specific",
                        "Larval",
                        "Increasing",
                        "L3_High",
                        "Not_Changing")

elt2_peaks$cluster.description <-
  factor(
    elt2_peaks$cluster.description,
    levels = c(
      "LE-specific",
      "Post-embryonic",
      "Increasing",
      "L3-high",
      "Not-changing or not IDR-passing"
    ),
    labels = elt2_cluster_names
  )

elt2_peaks %>% head
```
```{r}
table(elt2_peaks$cluster.description)
```

Merge the ELT-2 ChIP-seq data with the intestine filtered time-resolved RNA-seq data

```{r}
chip_rna_df <- elt2_peaks %>%
  dplyr::select(name, cluster.description, WBGeneID)  %>%
  right_join(time_resolved_rna,
             by = c("WBGeneID" = "wbps_gene_id")) %>%
  tidyr::replace_na(list("name" = "Not_Bound")) %>%
  filter(WBGeneID %in% intestine_gene_list$WBGeneID)
chip_rna_df$cluster.description <- factor(chip_rna_df$cluster.description, levels = c(levels(chip_rna_df$cluster.description), "Not_Bound"))
chip_rna_df$cluster.description[is.na(chip_rna_df$cluster.description)] <- "Not_Bound"
head(chip_rna_df)
table(chip_rna_df$cluster.description)
```


# Variance filter, log transform and z-score row normalize time-resolved RNA expression values

```{r chip_rna_matrix}
#### Handle duplicate rows created by 1:many gene:peak mapping

# match will return the first index of each non-redundant gene
nr_gene_name_ixs = match(unique(chip_rna_df$wikigene_name),chip_rna_df$wikigene_name)
# length(nr_gene_name_ixs)
#[1] 3012

# chip_rna_df = chip_rna_df[nr_gene_name_ixs,]

chip_rna_matrix <-
  chip_rna_df[nr_gene_name_ixs,] %>% dplyr::select(emb_510min:YA) %>% as.matrix()
rownames(chip_rna_matrix) <- chip_rna_df[nr_gene_name_ixs,]$wikigene_name
#### Handle 0's and take the log
# 1. Just replace 0's as NAs so we can apply log(). Alternatively, we could do log(x + .01), but there are only ~300
chip_rna_matrix_na = chip_rna_matrix; 
chip_rna_matrix_na[0 == chip_rna_matrix_na] <- NA
# 2. Apply log()
chip_rna_matrix_log = log( chip_rna_matrix_na )
# 3. Do variances row-wise, make sure to set na.rm=T
rowvariances = apply(chip_rna_matrix_log, 1, var, na.rm=T)

range(rowvariances) # no NaNs
# 4. Plot distribution of row variances of the log data...
# You can do hist with 10,100,1000 breaks, and there will always be
# a dominant spike all the way on the lowest value. This is because the data elicit no natural lowest break, and can be demonstrated by ecdf, which is binless.
plot(ecdf(rowvariances)) # no "steps" anywhere, just a smooth curve

# 5. Exclude the lowest 5% of the rows by their variance
# Remove NA's introduced by log transform
real = apply(chip_rna_matrix_log, 1, function(x) { ! any(is.na(x)) }) 
# Set threshold
thresh = 0.05
# Calculate row variance value at threshold
qthreshold = quantile(rowvariances,thresh)
# Identify matrix indexes where variance is above threshold
changing = rowvariances > qthreshold

# 6. Z score scale the log transformed values
# chip_rna_matrix = chip_rna_matrix_log
rownames(chip_rna_matrix_log) <- chip_rna_df[nr_gene_name_ixs,]$wikigene_name
chip_rna_matrix_scaled <- row_scale(chip_rna_matrix_log) # calls base::scale() via RWC23_Functions.R


```

# Time-resolved RNA heatmap based on ELT-2 occupancy cluster

```{r heatmap_function}
BoeckRNA_ELT2_chip_Heatmap <- 
  function(subsetrows, 
           column_title,
           name = "Boeck Time Resolved RNA",                                
           row_split=chip_rna_df[nr_gene_name_ixs,]$cluster.description[subsetrows],
           clustering_distance_rows = "euclidean",
           clustering_method_rows = "complete",
           ...)
           
{
  
  ix=which(rownames(chip_rna_matrix_scaled)[subsetrows] %in% gene_names)
  chip_GOI_df = data.frame(name=rownames(chip_rna_matrix_scaled)[subsetrows][ix], index=ix)
  
  BoeckRNA_ELT2_chip <- Heatmap(
  chip_rna_matrix_scaled[subsetrows,],
  name=name,
  row_split = row_split,
  column_title = column_title,
  row_title = NULL,
  cluster_columns = FALSE,
  clustering_distance_rows = clustering_distance_rows,
  clustering_method_rows = clustering_method_rows,
  ...
) +
  rowAnnotation(
    ELT2_cluster = chip_rna_df[nr_gene_name_ixs,]$cluster.description[subsetrows],
    col = list(
      ELT2_cluster = c(
        "Embryo_Specific" = "#7570B3",
        "Larval" = "#1B9E77",
        "Increasing" = "#E7298A",
        "L3_High" = "#D95F02",
        "Not_Changing" = "#505050",
        "Not_Bound" = "yellow"
      )
    ),
    border = TRUE
  ) +  rowAnnotation(foo = anno_mark(at = chip_GOI_df$index,
                               labels = chip_GOI_df$name))
BoeckRNA_ELT2_chip
}
```

```{r BoeckRNA_ELT2_chip}

gene_names <- c("bcl-11", "sre-23", "T25D10.4", "aat-6", "pgp-1", "ftn-1", "elt-2", "elt-7")


time_resolved_heatmap_ELT2_binding_split <- BoeckRNA_ELT2_chip_Heatmap(real&changing, sprintf("Boeck Time Resolved RNA\nRow Z-score of Log(RNA Timecourse)\nVariance excludes lower %.0f%% (n = %i)",
    thresh * 100, nrow(data.whole)))
time_resolved_heatmap_ELT2_binding_split

if (plot == TRUE){
  myPDFplot(time_resolved_heatmap_ELT2_binding_split, "TimeresolvedRNA_Heatmap_ELT2_Cluster_Split", width = 7, height = 6, plotdir)
}
```

# Cluster Time-resolved RNA-seq of intestine genes based on expression patterns

```{r BoeckRNA_ELT2_chip}
data.whole = chip_rna_matrix_scaled[real&changing,]
agglom='ward.D2'
hc = hclust(dist(data.whole), method =agglom )
# plot the hc height to get the screeplot
ggplot(NULL, aes(x=length(hc$height):1, y=hc$height)) +
    geom_point() + geom_line() +
    theme_bw() + labs(title="Scree Plot of hclust", x = "# of clusters", y="Height") + scale_x_continuous(limits=c(1,15)) 

nclust=3
# main 2 clusters are stable, 3rd unstable 
# set.seed(5)
# clusterboot( dist(data.whole), clustermethod=disthclustCBI,method=agglom,k=nclust)

```


```{r BoeckRNA_ELT2_chip, fig.width = 10, fig.height=6}
# 1. Record matrix indecies for genes of interest to label on heatmap
gene_names <- c("bcl-11", "sre-23", "T25D10.4", "aat-6", "pgp-1", "ftn-1", "elt-2", "elt-7")
subsetrows=real&changing
ix=which(rownames(chip_rna_matrix_scaled)[subsetrows] %in% gene_names)
chip_GOI_df = data.frame(name=rownames(chip_rna_matrix_scaled)[subsetrows][ix], index=ix)
# 2. Record ELT-2 occupancy pattern annotation per gene
boeck_chip_annotation <- as.data.frame.matrix(table(chip_rna_df$wikigene_name[real&changing], chip_rna_df$cluster.description[real&changing]))
boeck_chip_annotation <- make_cluster_annotation(data.whole, boeck_chip_annotation)
boeck_chip_annotation <- make_cluster_binary_annotation(boeck_chip_annotation)
# 3. Assign cluster names
rna_cluster_names <- c("Postembryonic", "Embryonic", "Late Postembryonic")
# 4. Draw the heatmap
time_resolved_heatmap <- Heatmap(
  data.whole,
  column_title = sprintf(
    "Boeck Time Resolved RNA\nRow Z-score of Log(RNA Timecourse)\nVariance excludes lower %.0f%% (n = %i)",
    thresh * 100, nrow(data.whole)
  ),
  name = "Row z-score log(dcpm)",
  cluster_columns = F,
  row_order = hc$order,
  row_split = cutree(hc, nclust),
  row_title = paste(rna_cluster_names, "\n(n = ",as.vector(table(cutree(hc, nclust))), ")", sep = ""),
  row_title_rot = 0,
  show_row_names = F
) +
  rowAnnotation(foo = anno_mark(at = chip_GOI_df$index,
                                labels = chip_GOI_df$name)) +
  binding_cluster_row_annotation(boeck_chip_annotation) +
  rowAnnotation(
    Not_Bound = boeck_chip_annotation$Not_Bound,
    col = list(Not_Bound = c(
      "absent" = "white", "present" = "black"
    )),
    border = TRUE
  )
time_resolved_heatmap

if (plot == TRUE){
myPDFplot(time_resolved_heatmap, "TimeresolvedRNA_Heatmap_ELT2_annotation", width = 10, height = 6, plotdir)
}

```

```{r BoeckRNA_ELT2_chip}
mosaic_df <- data.frame(Cluster = cutree(hc, nclust)) %>% rownames_to_column(var = "wikigene_name") %>% left_join(chip_rna_df %>% dplyr::select(wikigene_name, cluster.description), by = "wikigene_name")
mosaic_df$rna_cluster_name <- rna_cluster_names[mosaic_df$Cluster]
table(mosaic_df$rna_cluster_name, mosaic_df$cluster.description)
table(mosaic_df$Cluster)
tab <- table(mosaic_df$Cluster, mosaic_df$cluster.description)

hclusters = cutree(hc, nclust)
# tab=table(hclusters, chip_rna_df$cluster.description[real&changing])
dimnames(tab) <- list(
  "hclusters" = 1:nclust,
  "ELT-2" = levels(chip_rna_df$cluster.description)
)
tab
```


```{r custom-mosaic}
boeck_pearson_residuals <-
  as.data.frame(chisq.test(tab, simulate.p.value = TRUE)$stdres)
boeck_pearson_residuals <-
  cbind(boeck_pearson_residuals,
        as.data.frame(chisq.test(tab, simulate.p.value = TRUE)$observed)$Freq)
colnames(boeck_pearson_residuals) <-
  c("hclusters", "ELT2_cluster", "Pearson_Residual", "Observed")
boeck_pearson_residuals$hclusters <-
  factor(
    boeck_pearson_residuals$hclusters,
    levels = c("3", "2", "1"),
    labels = rna_cluster_names[3:1]
  )

# add q-value
q <- p.adjust(
  2*pnorm(
    abs(chisq.test(tab, simulate.p.value = TRUE)$stdres), 
    lower.tail = FALSE), 
  method = "BH")

tableformosiac.q <- matrix(q, ncol = 6)

dimnames(tableformosiac.q) <- dimnames(tab)
tableformosiac.q <- as.table(tableformosiac.q)
boeck_pearson_residuals <-
  cbind(boeck_pearson_residuals, 
        p.adjust = as.data.frame(tableformosiac.q)$Freq)
```

```{r fig.width= 5, fig.height=4}
# draw the plot
# Add pass vs fail for pvalue
ggplot(boeck_pearson_residuals,
       aes(x = ELT2_cluster, y = hclusters)) +
  geom_point(aes(size = Observed, fill = Pearson_Residual, colour = p.adjust < 0.05), shape = 21, stroke = 1) + 
  scale_fill_gradient2(
    high = "red",
    low = "blue",
    mid = "white",
    midpoint = 0
  ) +
  theme_bw() +
  theme(legend.key.width = unit(0.125,"in"), legend.key.height = unit(0.11, "in"), legend.key.size = unit(0.01, "in"), axis.text.x = element_text(angle = -45, hjust = 0)) +
  scale_color_manual(values = c("light grey", "black")) +
  xlab("ELT-2 Binding Cluster\n(from ChIP-seq)") +
  ylab("Time-resolved Gene Cluster\n(from RNA-seq)")

if (plot == TRUE){
ggsave(paste(plotdir,
          "TimeresolvedRNA_ELT2clusters_custom_mosaic",
          "_",
          today(),
          ".pdf",
          sep = ""), width = 5, height = 3.75, useDingbats=FALSE)
}


```